Analysis on Trend of Death of Shelter Residence in Toronto 

### 2024-09-24 paper/paper.qmd
add illustration of trend over month graph

# get line graph, showing the annual death trend
ggplot(annual_deaths, aes(x = year, y = total_yearly_deaths)) +
  geom_line(size = 1, color = "black") +  # get line graph
  geom_point(size = 3, color = "pink") +  # add points on value in each year
  labs(title = "Annual Deaths in Toronto Shelters (2007-2023)",
       x = "Year",
       y = "Total Deaths") +  # add title and name of the axis
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

### Compariasion Over Month

```{r fig-monthly-death, echo=FALSE, message=FALSE, warning=FALSE}

@@

# get line graph, showing the annual death trend
ggplot(month_deaths, aes(x = month, y = total_monthly_deaths, group = 1)) +
  geom_line(size = 1, color = "black") +  # get line graph
  geom_point(size = 3, color = "pink") +  # add points on value in each year
  labs(title = "Total Deaths in Toronto Shelters (2007-2023)",
       x = "Month",
       y = "Total Deaths") +  # as title and name of the axis
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

>>>>

# get line graph, showing the annual death trend
ggplot(annual_deaths, aes(x = year, y = total_yearly_deaths)) +
  geom_line(size = 1, color = "black") +  # get line graph
  geom_point(size = 3, color = "pink") +  # add points on value in each year
  labs(title = "Annual Deaths in Toronto Shelters (2007-2023)",
       x = "Year",
       y = "Total Deaths") +  # add title and name of the axis
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```

### Compariasion Over Month
Figure @fig-monthly-death shows the aggregated monthly death counts for shelter residents from 2007 to 2023. This data was derived by summing the death counts for each January from 2007 to 2023, resulting in a comprehensive data set for January, and repeating this process for each of the other months, resulting in data for all twelve months.

The purpose of this visualization is to analyze whether certain months have higher mortality rates. It opens the door to exploring potential seasonal influences on mortality rates, such as the impact of extreme temperatures or adverse weather conditions, which could contribute to increased mortality among shelter residents.

The current findings, shown in the line graph, indicate that January and November have the highest mortality rates, while March and September have the lowest mortality rates. These findings highlight the need for further research into the factors influencing these monthly variations, which could inform targeted interventions to improve the health and safety of shelter residents.

```{r fig-monthly-death, fig.cap="Monthly Mortality Trends Among Shelter Residents (2007-2023)", echo=FALSE, message=FALSE, warning=FALSE}

@@

# get line graph, showing the annual death trend
ggplot(month_deaths, aes(x = month, y = total_monthly_deaths, group = 1)) +
  geom_line(size = 1, color = "black") +  # get line graph
  geom_point(size = 3, color = "pink") +  # add points on value in each year
  labs(title = "Monthly Mortality Trends Among Shelter Residents (2007-2023)",
       x = "Month",
       y = "Total Deaths") +  # as title and name of the axis
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )
```
